We have derived expressions for the change in Von Neumann entropy due to a perturbation, including two beautiful and new expressions (\ref{eq.deltaS.noassumptions}) and (\ref{eq.Sredresult.separable}). We have seen that the leading-order perturbative correction to entropy is non-negative if the density operator's eigenvalue corrections for states in the initial density operator's kernel are of similar or lower order as for states not in the kernel. In particular, entropy will not decrease if we start in a separable mixed state such that some component states have zero initial probability but can be transitioned into. In this case, the entropy growth is dependent on the transition amplitudes to these states. We also applied our formulae in two toy examples.

Future work could examine whether it is possible to use the density of accessible kernel states to predict the rate at which systems become entangled. In particular, a system’s energy spectrum and selection rules might dictate the rate of decoherence.

In my analysis in Chapter \ref{ch.derivation}, it is possible that the first order-probability corrections \(\sigma_n^{(1)}\) vanish if and only if the initial state is separable. I have proven the ``if'' direction. The ``only if'' direction should also be proven or falsified by counterexample.

Future work should also investigate applications of entropy. For example, how useful is Von Neumann entropy for quantifying decoherence in quantum computing, and can the formulae derived here be used in such applications? 

Finally, physicists should continue to investigate how quantum entropy might give rise to thermodynamical properties. Perhaps one could derive all of thermodynamics starting with only the Schrödinger equation. Ref. \cite{bracken} and \cite{heusler} begin this work. It should be continued.
